---
title: "Identifying effective climate policy in Austria: a reverse-causal analysis"
author: "Talis Tebecis"
date: "`r Sys.Date()`"
output:
  pdf_document:
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(data.table)
pacman::p_load(dplyr)
pacman::p_load(tidyr)
pacman::p_load(readxl)
pacman::p_load(stringr)
pacman::p_load(gets)
pacman::p_load(getspanel)
pacman::p_load(here)
pacman::p_load(ggplot2)
pacman::p_load(ggpubr)
```

## Motivation

I seek to replicate the work of Koch et al. (2022) in *Attributing agnostically-detected large reductions in road* $CO_2$*emissions to policy mixes*, applying their 'causes-of-effects' searching method to determine effective emissions-reduction policy mixes across all sectors. I will then focus on the Austrian context to find effective policy mixes, in the context of informing the Climate Plan for Austria.

## Austrian emissions: an exploration

First, I visualise the overall emissions of Austria to explore potential structural breaks in emissions.

```{r, echo=FALSE, fig.height = 6, fig.width = 12}
emissions_codes <- read.csv("Data/01 Source_data/IPCC 2006 Categories.csv") %>% 
  select(IPCC, IPCC_description)

data_full <- read.csv("Data/02 Intermediary_data/CO2DriversEU_dataset_2022.csv") %>% 
  left_join(emissions_codes,
            by = c("category" = "IPCC")) %>% 
  filter(!is.na(lemissions_pc))

AUT_emissions <- data_full %>% 
  filter(country == "Austria") %>% 
  group_by(year) %>% 
  summarise(tot_emissions = sum(emissions))

AUT_emissions_pc <- data_full %>% 
  filter(country == "Austria") %>% 
  group_by(year) %>% 
  summarise(tot_emissions_pc = sum(emissions_pc))

plot1 <- ggplot(data = AUT_emissions) +
  geom_line(mapping = aes(x = year, y = tot_emissions))

plot2 <- ggplot(data = AUT_emissions_pc) +
  geom_line(mapping = aes(x = year, y = tot_emissions_pc))

ggarrange(plot1, plot2,
          labels = c("Total emissions for Austria",
                     "Total emissions per capita for Austria"),
          ncol = 2,
          nrow = 1)
```

Here, we can see significant variation in Austria's emissions, both upwards and downwards since 1970. In order to determine effective policy interventions to reduce emissions, we must consider emissions at the sectoral level.

\newpage

Thus, I explore the emissions per capita time series plots for all different IPCC Emissions categories. These are based on the 2006 IPCC emissions categorization guidelines.

```{r, echo=FALSE, fig.height = 6, fig.width = 12}
ggplot(data = subset(data_full, country == "Austria")) +
  geom_line(mapping = aes(x = year, y = emissions_pc)) +
  facet_wrap(~ IPCC_description) +
  labs(title = "Emissions per capita for Austria",
       xlab = "Year",
       ylab = "Emissions per capita") +
  theme(axis.text = element_text(size = 6))
```

We can see here that the IPCC categories with the highest emissions for Austria are:

-   Main Activity Electricity and Heat Production
-   Manufacturing Industries and Construction
-   Petroleum Refining - Manufacture of Solid Fuels and Other Energy Industries
-   Road Transportation
-   Other Sectors, which refers to Fuel Combustion Activities from the "Commercial/Institutional", "Residential" and "Agriculture/Forestry/Fishing/Fish Farms" sectors.

Next, I explore the variation in the time series by taking the log of per capita emissions and re-plotting the time series. I also allow the scale to vary between charts to see the individual variation in the plots.

```{r, echo=FALSE, fig.height = 6, fig.width = 12}
ggplot(data = subset(data_full, country == "Austria")) +
  geom_line(mapping = aes(x = year, y = emissions_pc)) +
  facet_wrap(~ IPCC_description, scale = "free") +
  labs(title = "Emissions per capita for Austria",
       xlab = "Year",
       ylab = "Emissions per capita") +
  theme(axis.text = element_text(size = 6))
```

From this analysis, we can see that there are significant variations in the emissions of all sectors. Thus, I conclude that it is worth evaluating structural breaks in emissions for all sectors for Austria. However, given the relatively large magnitude of the energy and heat, manufacturing and construction, refining, transport and other sectors, we should remember that emissions reductions in these sectors should be the focus, and not necessarily large relative emissions in low-emissions sectors.

### Level 1 analysis

```{r, echo=FALSE, fig.height = 6, fig.width = 12}
emissions_codes <- read.csv("Data/01 Source_data/IPCC 2006 Categories.csv") %>% 
  select(IPCC, IPCC_description)

data_full1 <- read.csv("Data/02 Intermediary_data/CO2DriversEU_dataset_2022_level1.csv") %>% 
  left_join(emissions_codes,
            by = c("category_name" = "IPCC_description")) %>% 
  filter(!is.na(lemissions_pc))

ggplot(data = subset(data_full1, country == "Austria")) +
  geom_line(mapping = aes(x = year, y = emissions_pc)) +
  facet_wrap(~ category_name, scale = "free") +
  labs(title = "Emissions per capita for Austria",
       xlab = "Year",
       ylab = "Emissions per capita") +
  theme(axis.text = element_text(size = 6))
```

### Level 2 analysis

```{r, echo=FALSE, fig.height = 6, fig.width = 12}
emissions_codes <- read.csv("Data/01 Source_data/IPCC 2006 Categories.csv") %>% 
  select(IPCC, IPCC_description)

data_full2 <- read.csv("Data/02 Intermediary_data/CO2DriversEU_dataset_2022_level2.csv") %>% 
  left_join(emissions_codes,
            by = c("category_name" = "IPCC_description")) %>% 
  filter(!is.na(lemissions_pc))

ggplot(data = subset(data_full2, country == "Austria")) +
  geom_line(mapping = aes(x = year, y = emissions_pc)) +
  facet_wrap(~ category_name, scale = "free") +
  labs(title = "Emissions per capita for Austria",
       xlab = "Year",
       ylab = "Emissions per capita") +
  theme(axis.text = element_text(size = 6))
```

## Identifying structural breaks

I follow the reverse causal approach of Koch et al. (2022) to identify structural breaks in the emissions per capita for each sector, which are not accounted for by changes in population or GDP per capita. I largely replicate the code of the original authors, but I have adjusted the dataset creation to account for all sectors, and have written a for-loop to iterate the process over all sectors and save the output accordingly.

I provide my code here for reference. Note, this takes a number of hours to execute. Also note, the "IPCC 2006 Categories" data is not in the original source code, and I have created this table based on the 2006 IPCC guidelines to make the categories more understandable (e.g. Road Transportation, instead of 1.A.3.b). The table also breaks the categories down into four layers of categorisation, potentially allowing for higher levels of granularity in assessing emissions reductions, for example, examining the Mineral Industry as a whole rather than cement, lime, glass and other carbonates separately.

The data and code are available at: <https://github.com/talistebecis/masters_thesis.git>.

```{r, eval = FALSE}
# Setup -------------------------------------------------------

rm(list=ls())
set.seed(1230)

#import data
data_full <- read.csv("Data/02 Intermediary_data/CO2DriversEU_dataset_2022.csv")

#Group specification
EU15 <- c("Austria", "Belgium", "Germany", "Denmark", "Spain", "Finland",
          "France", "United Kingdom", "Ireland", "Italy", "Luxembourg", 
          "Netherlands", "Greece", "Portugal", "Sweden")
EU16 <- c("Croatia", "Bulgaria", "Cyprus", "Czechia", "Estonia", 
          "Hungary", "Lithuania", "Latvia", "Malta", "Poland", "Romania", 
          "Slovak Republic", "Slovenia", "Switzerland", "Iceland", 
          "Norway")
EU31 <- c(EU15, EU16)

#IPCC categories
emissions_codes <- read.csv("Data/01 Source_data/IPCC 2006 Categories.csv") %>% 
  select(IPCC)

categories <- read.csv("Data/01 Source_data/IPCC 2006 Categories.csv") %>% 
  select(IPCC_description)


# Analysis ----------------------------------------------------------------

#run loop over all variables to produce results for all IPCC categories
for(IPCC_code in 15:24){
  
  #set IPCC code and category
  emissions_code <- emissions_codes[IPCC_code,]
  emissions_category <- categories[IPCC_code,]
  
  #set up document
  file_name <- paste0("Data/03 Output_data/2022/",emissions_category)
  
  #filter data
  data <- data_full %>% 
    filter(category == emissions_code) %>% 
    select(-category)
  
  #add lags
  data <- as.data.table(data)
  data[, L1.lemissions:=c(NA, lemissions[-.N]), by="country"]
  data[, L1.lgdp:=c(NA, lgdp[-.N]), by="country"]
  data[, L1.lpop:=c(NA, lpop[-.N]), by="country"]
  data[, L1.lemissions_pc:=c(NA, lemissions_pc[-.N]), by="country"]
  
  # Heterogenous effects preparation
  group.interactions <- c("lgdp_EU15", "lgdp_EU16", "lgdpsq_EU15", "lgdpsq_EU16",
                          "lpop_EU15", "lpop_EU16")
  data$lgdp_EU15 <- data$lgdp * (data$country %in% EU15)
  data$lgdp_EU16 <- data$lgdp * (data$country %in% EU16)
  data$lgdpsq_EU15 <- data$lgdp_sq * (data$country %in% EU15)
  data$lgdpsq_EU16 <- data$lgdp_sq * (data$country %in% EU16)
  data$lpop_EU15 <- data$lpop * (data$country %in% EU15)
  data$lpop_EU16 <- data$lpop * (data$country %in% EU16)
  
  #results document header
  cat(
    paste0(
      "#################################################################### \n",
      "#                                                                  # \n",
      "#                 CO2 DRIVERS EU - ANALYSIS                        # \n",
      "#                                                                  # \n",
      "#################################################################### \n",
      "\n \n \n"),
    file = file_name
  )
  
  # Analysis
  for(group in 1:2){
    
    # Prepare sample and data
    sample <- list(EU15, EU31)[[group]]
    dat <- filter(data, country %in% sample, year>=1995)
    
    # Print Sample Header
    cat(
      paste0(
        "############################## \n",
        "#  SAMPLE = EU", length(sample), " \n",
        "############################## \n",
        "\n \n "),
      file = file_name,
      append = T
    )
    
    for(p.value in c(.05, .01, .001)){
      
      # Break analysis:
      is <- isatpanel(
        data = dat,
        formula = ifelse(
          group == 1, "lemissions_pc ~ lgdp + lgdp_sq + lpop",
          paste0(
            "lemissions_pc ~ ", 
            paste(group.interactions, collapse = " + ")
          )
        ) %>% as.formula,
        index = c("country", "year"),
        effect = "twoways",
        iis = T,
        fesis = T, 
        t.pval=p.value
      )
      
      # Print analysis results
      cat(
        paste0(
          " \n ###########################", 
          " \n # p-value: ", p.value,
          " \n \n "), 
        file = file_name, 
        append = T)
      
      sink(file_name, append=T)
      print(is)
      sink()
      
      cat(" \n \n \n \n \n", 
          file = file_name, 
          append = T)
    }
  }
}
```

This allowed me to identify structural breaks for all country-year pairs in all IPCC Categories, except for Oil and Natural Gas, which produced the error, "Error in if (ar.LjungBox\$p.value \<= ar.LjungB[2]) { :missing value where TRUE/FALSE needed". As this was a relatively insignificant variable for Austria, I have not followed up on this yet.


## Structural breaks for Austria

I then collated the relevant results for Austria - everywhere a structural break was identified, whether it was positive or negative. This table shows the negative structural breaks, i.e. the emissions reductions not accounted for by reductions in GDP per capita or population.

```{r, echo=FALSE}
results <- read.csv("Data/03 Output_data/AUT_results_2022.csv")
negative_breaks <- results %>% 
  filter(coef <= 0) %>%
  select(category, sample, p.value.test, year, coef, Signif.) %>% 
  rename(p.value = p.value.test)
negative_breaks
```

We can see that only five negative breaks were detected for Austria across the two sample groups through this detection process:

-   2009: Incineration and Open Burning of Waste
-   2006: Lime Production
-   2006: Water-borne Navigation
-   2007: Water-borne Navigation
-   2015: Petroleum Refining - Manufacture of Solid Fuels and Other Energy Industries

Given the very close proximity of the Incineration and Open Burning of Waste and Water-borne Navigation breaks, it is likely that there are only 3 true structural breaks in the negative direction. I could do more testing here, using a dummy variable of more than 1 year for these breaks. I believe Koch et al. (2022) do something similar when they refer to Confidence Intervals, but it is not yet clear to me from their code or paper.

For completeness, I also include the positive structural breaks for Austria, of which there are 55, significantly more than the negative ones. These were omitted from the initial analysis of Koch et al. (2022), so I omit these from the attribution of policies for now.

```{r, echo=F}
positive_breaks <- results %>% 
  filter(coef > 0) %>%
  select(category, sample, p.value.test, year, coef, Signif.) %>% 
  rename(p.value = p.value.test)
positive_breaks
```

It is relevant to note that two of the IPCC Categories that we identified as being the largest emitters for Austria were included in the positive structural breaks:

-   Main Activity Electricity and Heat Production: 2003 and 2008
-   Road Transportation: 2001, 2002 and 2013

It could be worth analysing these upward breaks, perhaps to evaluate ineffective policies or unintended consequences from policies that could have caused the upward variation. For now, I have not pursued this line of thought.


## Attirbution to policies

I seek to attribute the negative structural breaks to policies or policy mixes as a means of testing policy effectiveness, using the following databases:

1.    **The IEA's Policies and Measures Database**: past, existing, or planned climate and energy policies. Data is collected from governments, international organisations, and IEA analyses, and governments can review the provided information periodically.
2.    **IEA/IRENA Renewable Energy Policies and Measures Database**: a joint database of renewable energy policies and measures of the IEA and IRENA.
3.    **The National Communications to the UNFCCC secretariat**: obligatory for our sample countries to submit regularly. Note, while this was used in the original paper, I am yet to see exactly how to identify individual policies from this database.
4.    **Other potential resources**: the World Bank's Carbon Pricing Dashboard that provides detailed information on carbon prices, and the Climate Change Laws of the World database of the Grantham Research Institute.

In the following, I try to attribute the negative structural breaks to policies implemented in those countries, and affecting the relevant sector, in the year of or years leading up to the break. Note, this is done on an ad hoc and manual basis, one structural break at a time. I have also only used resources (1) and (2) from the above list so far.


**Incineration of waste (2009)**

-   Ökostromverordnung - feed-in tariffs for green electricity (2009): Landfill gas, biomass and biogas, leading to waste being diverted from landfill. 15 year FiTs.
-   Klimastrategie - Climate Strategy (2008 - 2012)

**Lime production (2006)**

-   Emission Trading System implemented in 2005, which affected mineral industries
-   Expert System for an Intelligent Supply of Thermal Energy in Industry (EINSTEIN) (2007) is a methodology for the implementation of a holistic integral approach to thermal energy auditing in industry.

**Petroleum refining (2015)**

-   Residential building, energy and environmental subsidies (2014): reduced natural gas consumption

-   Mandatory Energy Audits for Large Companies (2015). This is a requirement for large companies to conduct mandatory energy audits. It first entered into force with the Energy Efficiency Act at the start of 2015, and the first audits were to be reported by the end of 2015.
-   Petroleum consumption and natural gas consumption reduced significantly in 2014
-   Only one oil refinery exists in Austria (Schwechat) - so overall emissions depend solely on one company (OMV)
-   Until the end of 2015, vehicles that run on alternative drivetrains (hybrids, those using fuels E85, CNG, LPG, or H2), receive a tax reduction of €600 ($639 US).
-   Since 2011, an increase in the mineral oil tax for conventional vehicles has been in effect — €0.04/L for gasoline and €0.05/L for diesel. As compensation for drivers, the commuting allowance was increased by 10%.
-   In 2014, Austria's public expenditures for energy-related R&D amounted to €143.1 million; an increase of 14.9% over expenditures in 2013 and representing an all-time high (Figure 5). The research areas of energy efficiency (43.1%), smart grids and storage (24.7%), and renewables (22.7%) define the priorities of publicly financed energy research within Austria. With €9.4 million in 2014, biofuels funding volume increased by €1 million in comparison to 2013.
-   Targets for renewables: 2007-10 Government programme.  Increase use of alternative fuels in the transport sector to 10% by 2010 and to 20% by 2020; development of a methane-based transport fuel with a share of at least 20% methane by 2010; provide fuel coverage with a network of E85 and methane filling stations by 2010.

**Water-borne navigation (2006-2007)**

-   klima:aktiv programme Renewable Energy (2005). Includes provisions for biogas and biomethane for transport use.

-   Klimaaktiv mobil (2007). Between 2007 and 2016 more than 6,600 green mobility projects have been initiated. This made annual savings of 610,000 tonnes of CO2 possible. In 2015,the BMLFUW allocated approx. 9 million EUR for klimaaktiv mobil funding andabout 2 million EUR for consulting, information and education programs, provided by the ministry, the Climate and Energy Fund and the National Environmental Support Scheme.


## Model

-   Multiple samples increases credibility for identified breaks
-   getspanel in R
-   Identifying structural breaks is commonplace in time-series literature on climate policy evaluation
-   Benefits: use of conservative confidence levels and use of control groups
-   “indicators”: interactions of country and year fixed effects - allow breaks in any country at any point in time
-   Countries are treated sparsely (machine learning)
-   EU15 are subject to largely identical EU regulations
-   Movement from general model to specific model using ML to the sparse model, using the block search algorithm "gets".



## Discussion with Jesus

My analysis is still preliminary and I have not conducted robustness checks. I have a number of open questions and/or actions that I need to take:

-   Should I aggregate industries at a higher level and repeat the analysis? Yes, to level 2 and 3
-   Is it worth looking at upward structural breaks rather than just negative? Nope, just down
-   How can I more systematically attribute policy mixes to breaks? This is poetry.
-   Can I improve the structural breaks estimation technique using Bayesian methods? This would be compared to the TWFE model with machine learning (the 'gets' algorithm for shrinkage) to make the model sparse. Yes, after the Masters Thesis
-   How do I replicate the confidence intervals around country-time pairs that the original authors used? Ask Anne and Lucas
-   How do I interpret the coefficients? What about values above 1.00 or below -1.00? Counterfactual
-   What other useful policy databases are there? Austrian government websites
-   Should I update to EDGAR v7 (currently v5) to get more recent emissions data up to 2021? Or will covid mess with the analysis? Yes

## Open questions and actions

-   Constructing confidence intervals
-   Counterfactual plots - done (somewhat)
-   What is the material benefit of the reductions detected? e.g. total annual tonnes of carbon saved
-   Which units to show? Analysis was done on log emissions per capita, but paper shows log emissions
-   Confirm gets package is using point and not step breaks (read Oxford GETS paper). Current model uses step changes

Model:
-   Read getspanel documentation
-   Read Pretis F, Schwarz M. Discovering What Mattered: Answering Reverse Causal Questions by Detecting Unknown Treatment Assignment and Timing as Breaks in Panel Models. mimeo. 2022.
-    Read: Pretis F. Does a carbon tax reduce CO2 emissions? Evidence from British Columbia. SSRN  Working Paper. 2019.

## PRESENTATION ACTIONS:

Policy - make into table
Conclusion - refine
Model - clean up and add formulas; understand gets and block search

